
### PART 2 : Stats

Part 1: Def

What is a median? A mean?

A median is obtain by adding values and dividing by the number of values. A median is the value in the middle of an ordered list of values.

<https://maths-et-tiques.fr/telech/Tstat1.pdf>

What is a variance? A standard deviation?

A variance indicates how spread out are the values from the mean value. It is calculated as the average of the squared differences between each value and the mean. The standard deviation is the square root of the variance, thus it is expressed in the same unit as the original values.

<https://maths-et-tiques.fr/telech/Tstat1.pdf>

What is a normal distribution?

A normal distribution has a bell shaped, symmetrical curve, 95% of the data lie within two standard derivations of the mean.

Najmi A, Sadasivam B, Ray A. How to choose and interpret a statistical test? An update for budding researchers. J Family Med Prim Care. 2021 Aug;10(8):2763-2767. doi: 10.4103/jfmpc.jfmpc_433_21. Epub 2021 Aug 27. PMID: 34660402; PMCID: PMC8483143.

What is a statistical test?

A statistical test is a method to determine whether there is enough evidence to state that there is a difference in what we want to test.

<https://www.itl.nist.gov/div898/handbook/prc/section1/prc13.htm?utm_source=chatgpt.com>

When to use a statistical test?

A statistical test is used when we want to determine if the difference observed in the data is real or due to chance.

<https://www.itl.nist.gov/div898/handbook/prc/section1/prc13.htm?utm_source=chatgpt.com>

What is a parametric test? A non-parametric test?

A parametric test is a type of test used when the data is normally distributed, whereas a non-parametric test is used when the data is not normally distributed.

Najmi A, Sadasivam B, Ray A. How to choose and interpret a statistical test? An update for budding researchers. J Family Med Prim Care. 2021 Aug;10(8):2763-2767. doi: 10.4103/jfmpc.jfmpc_433_21. Epub 2021 Aug 27. PMID: 34660402; PMCID: PMC8483143.

What are the assumptions of parametric tests?

The assumptions of parametric tests are that the data are normally distributed and that the variances are equal between groups.

<https://stats.libretexts.org/Bookshelves/Applied_Statistics/Mikes_Biostatistics_Book_%28Dohm%29/13%3A_Assumptions_of_Parametric_Tests?utm_source=chatgpt.com>

What are the assumptions of non-parametric tests?

The assumptions of non-parametric tests are that the sampling is random and they assume about the independence or dependence of samples.

<https://stats.libretexts.org/Bookshelves/Applied_Statistics/Statistics_for_Research_Students_%28Fein_Gilmour_Machin_and_Liam_Hendry%29/09%3A_Nonparametric_Statistics/9.01%3A_Nonparametric_Definitions?utm_source=chatgpt.com>

What is a p-value?

A p-value is a measure which state if a difference is statistically significant. More details to understand better: We start with a null hypothesis that there is no difference between what we want to test. We perform a statistical tests, we obtain a p-value.The p-value is the probability of obtaining a difference at least as large as the one observed, assuming that the null hypothesis is true. Hence, if the p-value is small, the observed result would be unlikely if the null hypothesis was true, so we reject the null hypothesis. Therefore, we state that there is a statistically significant difference (usually when p\<0.05).

<https://www.itl.nist.gov/div898/handbook/prc/section1/prc131.htm?utm_source=chatgpt.com>

What is the risk of error when using a statistical test?

The two types of errors are: - Type I error : when we reject the null hypothesis whereas the null hypothesis was true, - Type II error : when we fail to reject the null hypothesis whereas the null hypothesis was false.

<https://online.stat.psu.edu/stat500/Lesson06?utm_source=chatgpt.com>

What is the difference between a paired and an unpaired test?

A paired test is used when we test the same group multiple times. An unpaired test is used when testing the observations from two different groups.

<https://openbooks.library.unt.edu/quantitative-analysis-exss/chapter/statistical-evaluation-of-differnece/?utm_source=chatgpt.com>

# Part 2 : Effect of rehabilitation on performance

2.2.1

```{r}
# load the data and improt libraries
library(tidyverse)
library(dplyr)
PrePost <- read.csv("data/PrePost.csv")
head(PrePost)
summary(PrePost)
glimpse(PrePost)
  
```

| Variable | Description | Unit | Possible values | Expected change |
|---------------|---------------|---------------|---------------|---------------|
| perf | Performance measured during the rehabilitation assessment | Arbitrary units | 125-148 | Increase after rehabilitation |
| time | Measurement occasion | Category | Before, After | N/A |

## Does the treatment have an effect ? To answer, we will compare before and after

```{r noramlity test of the difference between before and after}

before <- PrePost$perf[PrePost$time == "Before"]
after  <- PrePost$perf[PrePost$time == "After"]

diff <- after - before # to check the normality of the difference between before and after, we will use the Shapiro-Wilk test

shapiro.test(diff)

```

## **What comparison are you going to make?**

- Which test will you use? Parametric or non-parametric?

  The compare performance before and after rehabilitation on the same participants, we first evaluated the normality of the paired differences, the Shapiro-Wilk test was not significant(p=0.283), confirming that the normality assumption was satisfied. Because the data are normally distributed and derived from a paired design, a parametric paired t-test is appropriate and was used for the comparison.

```{r performing the t-test on the before/after series}

t.test(after, before, paired = TRUE)

```

Interpretation : A paired t-test showed that performance was significantly higher after rehabilitation than before rehabilitation (t(7) = 5.40 from 0, p = 0.001). Participants improved by an average of 2.5 units (95% Coincidence Inteval \[1.41, 3.59\]). These results indicate that the rehabilitation program had a significant positive effect on performance.

- Which graph illustrates the effect of the treatment?

  These 2 plots are used to illustrate the effect of the tratement, the boxplot illustrates the distribution of the data and the second one illustrates the paired individual changes between the groups.

```{r Graph to illustrate effect over time}
#(before->after)
PrePost$time <- factor(PrePost$time, levels = c("Before", "After"))

#plot the 1st graph to visualize the distribution and the variation of before and after series.
ggplot(
  PrePost,
  aes(x = time, y = perf, fill = time)
) +
  geom_boxplot() +
  theme_minimal() +
  labs(
    x = "Time",
    y = "Performance",
    title = "Effect of rehabilitation on performance"
  )

PrePost$subject <- rep(1:8, 2)
#plot the 2nd graph to show the individual change due to rehabilitation.
ggplot(
  PrePost,
  aes(x = time, y = perf, group = subject)
) +
  geom_line(alpha = 0.6) +
  geom_point(size = 3) +
  theme_minimal() +
  labs(
    x = "Time",
    y = "Performance",
    title = "Individual changes in performance"
  )
```

- What sentence will you write in your thesis to explain your result?

The normality of the differences between pre- and post-treatment performance was first assessed using a Shapiro-Wilk test. The test was not significant (p = 0.283), indicating that the normality assumption was satisfied.

Consequently, a paired Student's t-test was performed to compare performance before and after rehabilitation. The analysis revealed a significant increase in performance after treatment (t(7) = 5.40, p = 0.001).

The mean improvement was 2.5 units (95% CI \[1.41, 3.59\]), indicating that participants performed significantly better following the rehabilitation program. These results suggest that the rehabilitation intervention had a positive effect on performance.

Limits

A major limitation of this analysis is the small sample size (n = 8). The results should therefore be interpreted with caution, as a larger sample would provide more precise and reliable estimates of the treatment effect.

Also,the study did not include a control group. Consequently, it is not possible to determine whether the observed improvement was exclusively due to the rehabilitation program or if others factors occured.
