
## 2.3.2 - The analysis part 2

```{r Import library and load data }

# Library
library(tidyverse)
library(dplyr)

# Import data
snore <- read.table("data/snore.txt", header = TRUE)

# Verify
head(snore)
str(snore)
```

## Are men fatter ?

To answer this question, we will first plot the data.

```{r Plot data }
#
ggplot(
  snore,
  aes(x = sex, y = weight, fill = sex)#fill #: we use fill to diference the two groups
) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(#some people same weight so shift a bit the points for more visability
    width = 0.15,
    alpha = 0.6
  ) +
  theme_minimal() +
  theme(
    legend.position = "none"#because sex is already indicated on the x axis
  ) +
  labs(
    title = "Body weight according to sex",
    x = "Sex",
    y = "Weight (kg)"
  )
```

Now we wil test if the data are normally distributed or not.

```{r Normality test }

shapiro.test(
  snore$weight[snore$sex == "H"]
)

shapiro.test(
  snore$weight[snore$sex == "F"]
)
```

The Shapiro-Wilk test showed that weight was not normally distributed in either group, with p=0.008 for men and p=0.039 for women.

Therefore, we will used Mann-Whitney-Wilcoxon because we want to compare two independent groups and because the data are not normally distributed.

```{r Mann-Whitney-Wilcoxon + descriptive statistics}

wilcox.test(
  weight ~ sex,#compare weight according to sex
  data = snore
)
snore %>% # %>% means take this result and give it to the next step
  group_by(sex) %>% #for each sex:
  summarise(
    mean_weight = mean(weight),
    median_weight = median(weight),
    sd_weight = sd(weight)
  )
```

The Mann-Whitney-Wilcoxon showed no significant difference in body weight between men and women (W=893.5, p=0.729).

The mean body weight was 90.77kg for men and 89.32kg for women.

Therefore, we found no statistically significant evidence that men were heavier than women in our sample.

## Do women smoke less?

First, we will create a table and a plot to have an idea of the data.

```{r Table and Plot }

#showing the % of smoking or not according to gender in a table
# N: Non smoker, O: Smoker
print("Table of smoker according to gender in percentage(%):")
smoking_table <- prop.table(
  table(snore$sex, snore$tabac),
  margin = 1 #calculate the proportion by line
) * 100 #to read it as percentage

rownames(smoking_table) <- c("Female", "Male")
colnames(smoking_table) <- c("No", "Yes")

print(smoking_table)

#graph 
ggplot(
  snore,
  aes(
    x = sex,
    fill = tabac
  )
) +
  geom_bar(position = "fill") +
  scale_x_discrete(
    labels = c(
      "F" = "Female",
      "H" = "Male"
    )
  ) +
  scale_fill_discrete(
    labels = c(
      "N" = "No",
      "O" = "Yes"
    )
  ) +
  
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  labs(
    title = "Smoking status according to sex",
    x = "Sex",
    y = "Proportion",
    fill = "Smoking"
  )
```

Because these two variables are categorical variables we use Chi-square test of independence to test whether sex and tabac are independent.

```{r Chi-square test}

chisq.test(table(snore$sex, snore$tabac))
```

The chi-square test showed a significant association between sex and smoking status (χ² = 7.0023, df = 1, p = 0.00814). In our sample, 40% of women were smokers compared with 72% of men. Therefore, the proportion of smokers was significantly lower among women than among men.

## From a more general perspective: Are there any correlations between variables?

We are interested in the quantitative variables:

- age

- weight

- height

- alcool

Therefore, we use Spearman's correlation to evaluate the correlations between these variables.

```{r Spearman-correlation}
num_data <- snore %>%
  select(age, weight, height, alcool)
#To detect correlation between numerical variables
cor(
  num_data,
  method = "spearman",
  use = "complete.obs"
)
```

The correlation analysis revealed very few relationships between the quantitative variables.

It showed one strong positive correlation between height and weight, with a Spearman correlation coefficient of 0.93. This means that taller individuals tended to be heavier and, heavier individuals tended to be taller.The other correlations were weak and close to zero.
