---
title: "Snore analysis Valentin"
author: "Valentin"
format: typst
editor: visual
---

```{r import data }
library(tidyverse)

snore <- read.table(
  "data/snore.txt",
  header = TRUE
)

glimpse(snore)
summary(snore)
```
```{r making factor from each category variable}
snore$snore <- factor(
  snore$snore,
  levels = c("N","O"),
  labels = c("No","Yes")
)

snore$sex <- factor(
  snore$sex,
  levels = c("F","H"),
  labels = c("Female","Male")
)

snore$tabac <- factor(
  snore$tabac,
  levels = c("N","O"),
  labels = c("No","Yes")
)

```
## Are snorers fatter?

```{r Answering the question if there's correlation between snorer and the weight}
#Checking if the series of data are follow the normality hypothesis

#check the snorer
shapiro.test(
  snore$weight[snore$snore=="Yes"]
)
#check the non-snorer one
shapiro.test(
  snore$weight[snore$snore=="No"]
)

```

Interpretation

The Shapiro-Wilk test revealed that weight was not normally distributed in either group (snorers: p = 0.024; non-snorers: p = 0.020).

Therefore, the assumption of normality was not satisfied then a non-parametric test was selected for the comparison.

```{r visualisation of the statistics }
ggplot(
  snore,
  aes(x = snore, y = weight, fill = snore)
) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(
    width = 0.15,
    size = 2,
    alpha = 0.7,
    color = "black"
  ) +
  scale_fill_manual(
    values = c(
      "No" = "steelblue",
      "Yes" = "tomato"
    )
  ) +
  labs(
    title = "Body weight of snorers and non-snorers",
    x = "Snoring status",
    y = "Weight (kg)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none"
  )

```

\`

```{r performing a non-parametric test : wilcox-test}
wilcox.test(weight ~ snore, data = snore)
```

```{r show the mean and median of weight in snorer&non snorer}
# compare the values between non-snorer & snorer in table
snore %>%
  group_by(snore) %>%
  summarise(
    mean_weight = mean(weight),
    median_weight = median(weight),
    sd_weight = sd(weight)
  )
```

Discussion

To determine whether snorers were heavier than non-snorers, the normality of weight distributions was assessed using the Shapiro-Wilk test. Since normality was rejected in both groups (p \< 0.05), a Mann-Whitney-Wilcoxon test was performed.

The analysis revealed no significant difference in body weight between snorers and non-snorers (W = 1154, p = 0.908). The graph coincide with previous results.

